---
title: "Lab 01 - Hello R"
author: "Mia Williams"
date: "01/17/2024"
output: 
  html_document: 
    fig_height: 7
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
datasaurus_dozen <- datasaurus_dozen %>% 
  rename(shape = dataset)
```

If you've just finished running the first `load-packages` chunk above, you can navigate back to your Lab 01 instructions now.

## Exercises

### Exercise 1

```{r ?datasaurus_dozen}
# replace this line with ?datasaurus_dozen
?datasaurus_dozen
```

3 columns, 1846 rows
variables: dataset, x-values, y-values

### Exercise 2

First let's plot the data in the `dino` shape:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(shape == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

I copied and pasted the code from the dinosaur plot earlier and replaced "dino" with "plot-star". I also had to delete some extra text that I determined wasn't necessary through trial and error running (it told me what I didn't need after I ran it). 
no first person, concise
```{r plot-star,fig.height=3, fig.width=6}
star_data <- datasaurus_dozen %>%
  filter(shape == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()

star_data %>%
  summarize(r = cor(x, y))

```

Between dino and star, the correlation coefficients are NOT the same but they are similar. Dino is -0.065 and star is -0.063. 

```{r cor-star, fig.height=3, fig.width=6}
star2_data <- datasaurus_dozen %>%
  filter(shape == "star")

ggplot(data = star2_data, mapping = aes(x = x, y = y)) +
  geom_point()

```



### Exercise 4

Add code and narrative below as needed.  -1

Note that two R chunks are given but they are not labeled.  

Use the convention from above to name them appropriately.

```{r plot-circle,fig.height=3, fig.width=6}
circle_data <- datasaurus_dozen %>%
  filter(shape == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

circle_data %>%
  summarize(r = cor(x, y))

```

The circle's coeffeicient is -0.068, slightly smaller than dino's. 

```{r}

```

### Exercise 5
-.5
```{r}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = shape))+
  geom_point()+
  facet_wrap(~ shape, ncol = 3) +
  theme(legend.position = "none")
```

```{r}
datasaurus_dozen %>%
  group_by(shape) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```
  

**The fastest way to insert a code chunk is to use the keyboard shortcut: command-option-i (on MacOS) or Control-Alt-i (on Windows).** Or you can add R chunks by manually typing out the backticks, curly braces, and the letter `r` or use the Insert chunk button above, green C+.

-1 repo name should be unit-1-williams-mia or unit-1-williams
